---
title: 'Task 2: Survey responses - political and environmentalist self-identification'
author: "Katie Pietrucha"
date: "12/01/2021"
output:
  html_document:
    theme: yeti
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(here)
library(tidyverse)
library(janitor)
library(kableExtra)
library(broom)
```

## Introduction

This report utilizes the results of a survey conducted by Chelsea Batavia and Andrews Forest LTER to explore "Influences on charitable giving for conservation." The survey was conducted in August 2017 and yielded 1,331 individual respondents. The following analysis explores the association between the respondents’ self-identified political affiliation (conservative or liberal) and if they consider themselves an environmentalist. A chi-square test is used to determine if a significant association exists.

```{r, include = FALSE}
# Read in the data

conservation_survey <- read_csv(here("data", "conservation_survey.csv")) %>%
  clean_names()

# Wrangle to keep only env_id and pol_or variables

survey_subset <- conservation_survey %>%
  select("env_id", "pol_or")
```

```{r, include = FALSE}
# Add new column to bin political orientation into a binary response of conservative or liberal.

survey_subset_full <- survey_subset %>%
  mutate(pol_or_full = case_when(pol_or %in% c(1:3) ~ "Conservative",
          pol_or %in% c(5:7) ~ "Liberal")) %>%
          na.omit()
```

## Results 

This table shows the counts and proportions of “liberal” and “conservative” survey respondents who responded “Yes”, “No” and “Unsure” to the statement “I consider myself an environmentalist.” 

```{r, include = TRUE}
# Create finalized table

survey_counts <- survey_subset_full %>% 
  tabyl(pol_or_full, env_id) 

survey_proportions <- survey_counts %>%
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()

  colnames(survey_proportions) = c(" ", "Yes", "No", "Unsure")

survey_table <- survey_proportions %>%
  kbl(caption = "<center><strong>Table 1:</strong> Counts and proportions of responses to the statement 'I consider myself an environmentalist,' divided by self-identified political orientation (Batavia 2019).</center>") %>%
  kable_classic_2(full_width = F) 

survey_table

```

```{r, include = FALSE}
# Create contingency table

survey_ct <- survey_counts %>% 
  column_to_rownames(var = "pol_or_full")
```

```{r, include = FALSE}
# Run chi-square

survey_x2 <- chisq.test(survey_ct)
survey_x2

survey_tidy <- tidy(survey_x2)
```

## Analysis

From the counts and proportions, one can see that a larger number and proportion (318, 50.24%) of self-identified conservatives declined to also identify as environmentalists. On the other hands, more liberals identified as environmentalists (281, 53.73%). However, the results of the chi-square show that there is no significant association between political orientation and environmentalist self-identification ($\chi$^2^(`r survey_tidy$parameter`)) = `r round(survey_tidy$statistic, 2)`, *p* < 0.001) in these survey results. Political orientation and environmentalist self-identification are independent of each other. 

## Data Citation

Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c
