---
title: 'Task 3: Female Lizard Lengths'
author: "Katie Pietrucha"
date: "12/01/2021"
output:
  html_document:
    theme: yeti
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(paletteer)
library(kableExtra)
library(effsize)
library(broom)
```

## Introduction

This project explores lizard pitfall trap data from the Jornada Basin LTER in New Mexico. The analysis investigates the difference in total body length between female lizards of two species -- the Colorado Checkered lizard (*Cnemidophorus tessalatus*) and the Desert Grassland lizard (*Cnemidophorus uniparens*) -- using a two-sample t-test and Cohen's d effect size.

```{r, include = FALSE}
# Read in the data

lizards <- read_csv(here("data", "lizards.csv"))

# Simplify data frame, selecting variables for species, sex, and total length, and filtering for female CNUN and CNTE lizard observations.

lizards_subset <- lizards %>%
  select(spp, sex, total_length) %>%
  filter(sex == "F", spp %in% c("CNUN", "CNTE"))
```

## Results

```{r}
# Finalized visualization comparing total lengths of female lizards for the two species.

lizards_subset %>%
  ggplot(aes(x = total_length, y = spp)) +
  geom_boxplot(lwd = 0.5, aes(color = spp)) +
  geom_jitter(aes(color = spp), alpha = 0.5, cex = 2) +
  theme_bw() +
  labs(x = "Total length (mm)", y = "Species", caption = "Figure 1: Comparison of total length (mm) between females of two New Mexico lizard species. Each point is a single 
observation. Box endpoints indicate the 25th and 75th percentile values; the bolded line within each box indicates 
the mean value for each species. Mean length is higher for the Colorado Checkered lizard (C. tessalatus) than the 
Desert Grassland lizard (C. uniparens) (Lightfoot 2017).") +
  scale_color_manual(values = c("lightsteelblue4", "olivedrab4"), name = "Lizard Species", labels = c("Colorado Checkered
  (C. tessalatus)", 
  "Desert Grassland
  (C. uniparens)")) +
  theme(legend.title.align = 0.5) +
  theme(axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()) +
  theme(legend.position = "bottom") +
  theme(legend.text = element_text(face = "italic")) +
  theme(plot.caption = element_text(hjust = 0))
```

```{r}
# Finalized summary table with mean, standard deviation, and sample size.

lizards_stats <- lizards_subset %>%
  group_by(spp) %>%
  drop_na() %>%
  summarize(mean = round(mean(total_length), 2),
            sd = round(sd(total_length), 2),
            n = n()) 

lizards_table <- lizards_stats %>%
  mutate(spp = c("Colorado Checkered
  (C. tessalatus)", "Desert Grassland
  (C. uniparens)")) %>%
  kbl(col.names = c("Lizard Species", "Mean Total Length (mm)", "Standard Deviation (mm)", "Sample Size"), caption = "<center><strong>Table 1:</strong> Mean total length (mm), standard deviation (mm) and sample size for female lizards of two species (Lightfoot 2017).</center>") %>%
  kable_classic(full_width = F, position = "left")

lizards_table
```

```{r, include = FALSE}

# Exploratory histogram
ggplot(data = lizards_subset) +
  geom_histogram(aes(x = total_length))

# Exploratory q-q plot
ggplot(data = lizards_subset) +
  geom_qq(aes(sample = total_length))

#Distribution looks fairly normal
```

```{r, include = FALSE}
# T-test
cnun_sample <- lizards_subset %>% 
  filter(spp == "CNUN") %>% 
  drop_na() %>%
  pull(total_length) 

cnte_sample <- lizards_subset %>% 
  filter(spp == "CNTE") %>% 
  drop_na() %>%
  pull(total_length)

lizard_t <- t.test(cnte_sample, cnun_sample)

t_tidy <- tidy(lizard_t)
```

```{r, include = FALSE}
# Find effect size (Cohen's d)

lizards_cohen <- cohen.d(cnte_sample, cnun_sample)

lizards_cohen
```
## Analysis

The mean total length for the female Colorado Checkered lizard was `r lizards_stats$mean[1]` +/- `r lizards_stats$sd[1]` mm, and the mean total length for the female Desert Grassland lizard was `r lizards_stats$mean[2]` +/- `r lizards_stats$sd[2]`  mm. The effect size is large (`r round(lizards_cohen$estimate, 2)`) and the difference significant (two-sample t-test: t = `r round(t_tidy[4], 2)`, p < 0.001).

## Data Citation

Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative. https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e 
